Personality and political beliefs across the lifespan

### 2019-11-20 scripts/manuscript.Rmd
figure captions, recoding variables

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Someone should write an abstract

@@

```{r tidying}
#last step is to tidy our data
tidy <- data %>% 
  pivot_longer(cols = c(sec_economic_conservatism, sec_social_conservatism),
               names_to = "conservatism_type",
               names_prefix = "sec_",
               values_to = "conservatism_score") %>% 
  pivot_longer(cols = agreeableness:emotional_stability,
               names_to = "personality_trait",
               values_to = "personality_score")
```

```{r figure1}
ggplot(tidy, aes(x = age, y = personality_score, color = personality_trait)) +
  geom_smooth(se=FALSE, method = "lm")+
  labs(title = "Personality traits across the lifespan", subtitle = "Big 5 traits self-reported by adults aged 20-68",
       x= "Age", y = "How strongly participants identify with the personality trait", 
       color = "Personality Trait")+
  theme_classic()+
  scale_color_manual(values = c("red", "orange", "yellow","green", "blue"),
                     labels=c("Agreeableness", "Conscientiousness", 
                              "Emotional Stability", "Extraversion",
                             "Openness to experiences"))

@@

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

>>>>

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Here is where we will write our abstract.

@@

```{r tidying}
#last step is to tidy our data
tidy <- data %>% 
  pivot_longer(cols = c(sec_economic_conservatism, sec_social_conservatism),
               names_to = "conservatism_type",
               names_prefix = "sec_",
               values_to = "conservatism_score") %>% 
  mutate(conservatism_type = recode(conservatism_type, `economic_conservatism` = "Economic conservatism",
                                    `social_conservatism` = "Social conservatism")) %>%
  pivot_longer(cols = agreeableness:emotional_stability,
               names_to = "personality_trait",
               values_to = "personality_score") %>% 
  mutate(personality_trait = recode(personality_trait, `agreeableness` = "Agreeableness",
                                    `conscientiousness` = "Conscientiousness",
                                    `openness_to_experiences` = "Openness to experiences",
                                    `extraversion` = "Extraversion",
                                    `emotional_stability` = "Emotional stability"))
```

@@

```{r figure1, fig.cap= "Generally, older adults rated themselves higher on each of the big five personality traits than young adults. This sample was unexpectedly introverted."}

ggplot(tidy, aes(x = age, y = personality_score, color = personality_trait)) +
  geom_smooth(se=FALSE, method = "lm")+
  labs(title = "Personality traits across the lifespan", subtitle = "Big 5 traits self-reported by adults aged 20-68",
       x= "Age", y = "How strongly participants identify with the personality trait", 
       color = "Personality Trait")+
  theme_minimal()+
  scale_color_viridis_d()

```{r figure2, fig.cap= "Generally, older adults reported stronger socially and fiscally conservative beliefs compared to young adults. This increase in conservative beliefs as we age may be more extreme for socially conservative beliefs than for fiscally conservate beliefs."}

ggplot(tidy, aes(x = age, y = conservatism_score, color = conservatism_type)) +
  geom_smooth(se=FALSE, method = "lm")+
  scale_color_viridis_d() +
  labs(title = "Conservatism across the lifespan (adults aged 20-68)", subtitle = "Difference in Economic and Social Conservatism by Gender",
       x= "Age", y = "Conservatism Score (higher = more conservative)",
       color = "Conservatism Type")+
  theme_minimal()+
  facet_wrap(~gender)
```

```{r messy_data}
messy <- tidy %>% 
  pivot_wider(
    names_from = conservatism_type, 
    values_from = conservatism_score) %>%
   pivot_wider(
     names_from = personality_trait,
     values_from = personality_score)

#we can use this messy data to run lm models
```

# Methods

### 2019-11-20 scripts/manuscript.tex
figure captions, recoding variables

\abstract{
Someone should write an abstract

@@

\includegraphics{manuscript_files/figure-latex/figure1-1.pdf} \# Methods
We report how we determined our sample size, all data exclusions (if
any), all manipulations, and all measures in the study.

>>>>

\abstract{
Here is where we will write our abstract.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table1}Mean Personality Trait Scores Across the Lifespan}
\begin{tabular}{llllll}
\toprule
 & \multicolumn{5}{c}{Age Group} \\
\cmidrule(r){2-6}
Personality Trait & \multicolumn{1}{c}{[20-29]} & \multicolumn{1}{c}{[30-39]} & \multicolumn{1}{c}{[40-49]} & \multicolumn{1}{c}{[50-59]} & \multicolumn{1}{c}{[60-68]}\\
\midrule
Agreeableness & 4.70 & 5.11 & 5.58 & 6.05 & 6.08\\
Conscientiousness & 5.24 & 5.60 & 5.89 & 5.71 & 5.88\\
Emotional stability & 4.22 & 4.91 & 4.97 & 5.79 & 5.71\\
Extraversion & 3.14 & 3.42 & 3.37 & 4.18 & 3.96\\
Openness to experiences & 5.10 & 5.22 & 4.97 & 5.50 & 6.04\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Personality trait scores reported on a 1-7 scale.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/figure1-1.pdf}
\caption{\label{fig:figure1}Generally, older adults rated themselves higher
on each of the big five personality traits than young adults. This
sample was unexpectedly introverted.}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/figure2-1.pdf}
\caption{\label{fig:figure2}Generally, older adults reported stronger
socially and fiscally conservative beliefs compared to young adults.
This increase in conservative beliefs as we age may be more extreme for
socially conservative beliefs than for fiscally conservate beliefs.}
\end{figure}

\section{Methods}\label{methods}
